\ttitle

### 2015-08-08 Chapters/Introduction.tex
Introduction - Obesity and Survival

\section{Body Composition}
\todo{Introduction - body composition}

Obesity is an increasing problem in all developed nations with the proportion of obese patients undergoing surgery increasing every year. 40\% of the adult population in the United Kingdom are expected to have a body mass index of over 30 kg/m$^2$ by 2025. Obesity, especially morbid obesity, is associated with a wide range of secondary disease states such as ischaemic heart disease, diabetes mellitus, hypertension and increased incidence of cancers.

@@

\subsection{Obesity and Surgical Outcomes}

@@

Several studies have reported increased incidence of postoperative complications after cardiac, transplant, urologic and gynaecological surgery. \parencite{prem_operative_1965, holley_obesity_1990, fasol_influence_1992, gruberg_impact_2002, lee_impact_2004} Visceral obesity rather than body mass index has been reported to be associated with greater incidence of wound complications, overall complication rate and prolonged hospital stay in pateints undergoing laparascopic sigmoid colectomy for cancer. \parencite{tsujinaka_visceral_2008}

However, in a prospective, multi-institutional, risk-adjusted cohort study of 2258 patients including 699 patients who underwent pancreatectomy for cancer, Mullen and co-workers reported that obesity on its own was not a predictor of major postoperative complications or mortality. However, they noted that underweight patients had a five-fold increased risk of postoperative mortality. They also reported that obesity was an independent predictor of minor complications including wound infections.\parencite{mullen_impact_2008}

Dindo and co-workers studied the impact of obesity on postoperative outcomes in a prospective study of 6336 patients undergoing elective general surgery. They reported that obesity was not an independent risk factor for morbidity or mortality. There was no difference in the median operating time, blood transfusions or lenght of hospital stay between obese and non-obese patients. The type and severity of complications were similar between obese and non-obese patients. \parencite{dindo_obesity_2003}

In a more recent prospective, multi-institutional, risk-adjusted cohort study of 118,707 patients undergoing non-bariatric general surgery, Mullen co-workers reported that the risk of mortality was lowest in the overweight (BMI 25 - 30) and moderately obese (BMI 30 - 35) patients and highest in the underweight and morbidly obese patients. This appears to confirm the existence of an 'obesity paradox' in surgical patients. \parencite{mullen_obesity_2009}

>>>>

\section{Body Composition, Obesity and Outcomes}
\todo{Introduction - body composition}

The Body Mass Index (BMI) is a measure of a person's weight relative to their height and corresponds well with their body composition, especially with the proportion that is body fat. It is the most widely accepted measure of body composition, especially of body fat in spite of criticisms that it may not apply to all populations. BMI was first devised by Adolphe Quetelet and was formerly known as the Quetelet index.\parencite{eknoyan_adolphe_2008} Quetelet was a Belgian sociologist and mathematician who played a major role in the application of statistical methods to the social sciences. In his attempts to describe the characteristics of 'normal man', Quetelet found that weight increased nearly as the square of the height except at times of growth spurts in the new born and during puberty. \parencite{quetelet_treatise_2013} The calculation of body mass index is simple and as follows:

\begin{equation} \label{eq:bmi}
Body\ Mass\ Index = \frac{Weight\ in\ Kilograms}{(Height\ in\ Metres)^2}
\end{equation}

The World Health Organisation uses thresholds to categorise adult populations based on body mass index. These thresholds are independent of age, gender and race. However, there has been increasing criticism of these thresholds and the World Health Organisation is undertaking a further review to assess the validity of these thresholds in different populations. For instance, there is increasing evidence in Asians, the risk of diabetes and cardiovascular disease is greater at BMI levels lower than the current WHO cut-off for being overweight. Nonetheless, the thresholds shown in Table \ref{table:bmi_who} are currently in universal use and have been used in this thesis.

\begin{table}[h]
	\centering
	\caption{The WHO classification of Body Mass Index}
	\label{table:bmi_who}
	\begin{tabular}{ |l|l|}
		\hline
		BMI         & Category                  \\ \hline
		Below 18.5  & Underweight               \\
		18.5 – 24.9 & Normal range              \\
		25.0 – 29.9 & Overweight or Pre-obesity \\
		30.0 - 34.9 & Obesity class 1             \\
		35.0 - 39.9 & Obesity class 2             \\
		Above 40    & Obesity class 3             \\ \hline
	\end{tabular}
\end{table}

Obesity is an increasing problem in all developed nations with the proportion of obese patients undergoing surgery increasing every year. Among the European Member States, the proportion of adults  considered overweight or obese varied in 2008 between  37.0\% and 56.7\% for women and 51.0\% and 69.3\% for men. The United Kingdom had the highest proportion of obese women (23.9\%) and the second highest proportion of obese men (22.1\%). A greater proportion of men than women were obese and there was a marked increase in BMI with increasing age. However, as the proportion of overweight or obese decreased in women as their education level increased.

Obesity, especially morbid obesity, is associated with a wide range of secondary disease states such as ischaemic heart disease, diabetes mellitus, hypertension and increased incidence of cancers.

@@

\subsection{Obesity and Postoperative Complications}

@@

Several studies have reported increased incidence of postoperative complications after cardiac, transplant, urologic and gynaecological surgery. \parencite{prem_operative_1965, holley_obesity_1990, fasol_influence_1992, gruberg_impact_2002, lee_impact_2004} Visceral obesity rather than body mass index has been reported to be associated with greater incidence of wound complications, overall complication rate and prolonged hospital stay in patients undergoing laparascopic sigmoid colectomy for cancer. \parencite{tsujinaka_visceral_2008}

However, in a prospective, multi-institutional, risk-adjusted cohort study of 2258 patients including 699 patients who underwent pancreatectomy for cancer, Mullen and co-workers reported that obesity on its own was not a predictor of major postoperative complications or mortality. But,obesity was an independent predictor of minor complications including wound infections. However, they noted that underweight patients had a five-fold increased risk of postoperative mortality. \parencite{mullen_impact_2008}

Dindo and co-workers studied the impact of obesity on postoperative outcomes in a prospective study of 6336 patients undergoing elective general surgery. They reported that obesity was not an independent risk factor for morbidity or mortality. There was no difference in the median operating time, blood transfusions or length of hospital stay between obese and non-obese patients. The type and severity of complications were similar between obese and non-obese patients. \parencite{dindo_obesity_2003} In a more recent prospective, multi-institutional, risk-adjusted cohort study of 118,707 patients undergoing non-bariatric general surgery, Mullen and co-workers reported that the risk of mortality was lowest in the overweight (BMI 25 - 30) and moderately obese (BMI 30 - 35) patients and highest in the underweight and morbidly obese patients. This appears to confirm the existence of an 'obesity paradox' in surgical patients. \parencite{mullen_obesity_2009}

In a study of 356 consecutive patients undergoing pancreaticoduodenectomy at the Memorial Sloan-Kettering Cancer Center between 2000 and 2005, House and co-workers used preoperative cross-sectional imaging of the abdomen (computer tomography and/or magnetic resonance imaging) to measure a variety of indices of body composition including abdominal wall fat thickness, hip girdle fat thickness, retro-renal visceral fat and pancreatic duct diameter. \parencite{house_preoperative_2008} The incidence of wound infection was greater in obese patients with the risk increasing linearly with BMI. But, BMI was not associated with any other complications. They also reported that the thickness of retro-renal fat, a surrogate for visceral adipose tissue, was a significant risk factor for pancreatic fistula as well as wound infection. However, there was no correlation between complications and abdominal wall thickness or hip girdle fat. The importance of visceral fat in the incidence and severity of postoperative complications is further emphasised by the finding that excess visceral fat is associated with the the occurence of pancreatic fistula after radical total gastrectomy for gastric cancer. Benns and co-workers reported greater blood loss and increased incidence and severity of postoperative complications in obese pateints undergoing surgery compared to non-obese pateints. However, there was no difference in operative time or length of stay in hospital. \parencite{benns_impact_2009}

On the contrary, other authors have reported that obesity is not associated with increased complications after pancreatic surgery. Khan and co-workers reported that while obesity was associated with increased incidence of intra-operative bleeding, there was no increase in complications. \parencite{khan_does_2010} Length of stay in hospital, overall morbidity and in-hospital mortality were not affected by BMI after pancreaticoduodenectomy \parencite{tsai_impact_2010} while neither BMI nor intra-abdominal visceral fat affected postoperative outcomes in another study of patients undergoing pancreaticoduodenectomy. \parencite{balentine_obesity_2011}

In a recent meta-analysis of 17 studies, including some of those discussed above, Ramsey and co-workers reported that BMI was not associated with length of hospital stay, in-hospital mortality  or survival after pancreaticoduodenectomy for cancer. Obesity was associated with postoperative pancreatic fistula. However, the authors concluded that with aggressive perioperative care, postoperative morbidity in obese patients could be mitigated. They also noted several limitations to making meaningful comparisons between studies due to the lack of clear definitions for complications after pancreaticoduodenectomy. \parencite{ramsey_body_2011}

%lin_risk_2004, lermite_risk_2007 - Coranary artery disease is associated with pancreatic fistula
\subsection{Obesity and Cancer Survival}

The effect of obesity and adipose tissue content on cancer-specific survival in patients with pancreatic cancer is not entirely clear. In a study involving 586 patients who underwent pancreaticoduodenectomy for pancreatic adenocarcinoma, Khan and co-workers reported that obesity was not associated with overall or disease-free survival. \parencite{khan_does_2010} These are similar to the findings reported by Benns and co-workers who found no difference in overall survival in 306 patients undergoing resections for pancreatic adenocarcinoma. \parencite{benns_impact_2009} Obesity during early adulthood is associated with increased risk of pancreatic cancer while obesity at an older age is associated with reduced overall survival in both resected and unresected patients. \parencite{li_body_2009} Morbidly obese patients (BMI $>$ 35) were found to be at 12-fold increased risk of lymph node metastases and associated decreased survival. \parencite{fleming_influence_2009}

Tsai and co-workers reported that while obese patients had similar tumour characteristics to non-obese patients and comparable perioperative outomes, long-term survival was better in obese subjects. \parencite{tsai_impact_2010}  This study of 795 patients undergoing potentially curative surgery for pancreatic adenocarcinoma at the Johns Hopkins Medical Institutions in the United States is in contrast to other similar studies. The role of obesity on long-term survival after pancreaticoduodenectomy for pancreatic cancer is less clear with more recent studies suggesting that there was no link between body mass index and survival. \parencite{dandona_influence_2011, benns_impact_2009}
